State and Parameter Estimation for Spatio-Temporal Neural Fields

### 2010-05-10 ltx/Journal of Neuroengineering/StateAndParameterEstimationNeuralFields_JoNE.tex
plot of Mean of absolute error and the std
editting paper body

where $\mathbf{\boldsymbol{\phi}}(\mathbf{r})$ is a vector of Gaussian basis functions that are scaled by the state vector, $\mathbf{x}_t$. The choice of Gaussian basis functions can be justified by the existence of the so called bump solutions for this class of model, which have a Gaussian shape~\cite{Coombes2005}. The width and positioning of the basis function can be determined by spectral analysis (explained in detail in Section~\ref{SpectralAnalysisSection}). The connectivity kernel can also be considered a decomposition into a set of basis functions in 

@@

There are two options available to alleviate problems with aliasing. The first, which is the most desirable, is to increase the spatial sampling frequency (using a higher resolution of sensors). The second is to use wider sensors, where the spatial averaging acts like a low-pass anti-aliasing filter. This option is a compromise between alleviating issues with aliasing and loosing high frequency information contained in the field. In practice, it is difficult  to estimate the bandwidth of the cortex using traditional electrophysiological measurements due to spatial limitations. However, we envisage it may be possible the estimate it using other modalities with higher spatial resolution such as, fMRI, NIRS or other optical imaging techniques~\cite{Issa2000}. Errors associated with mild aliasing result in distortions in the high frequency components of the field dynamics. However, after applying the basis function decomposition, high frequency components contribute negligibly to the inverse problem and are therefore not a major problem for the state and parameter estimation~\cite{Sanner1992}. This is further illustrated our results. \omg{Although our results are severely effected by the loss of high frequency info! Ken: Do you think you can word this a little better than what I have done.}

Similar considerations need to be made regarding the representation of the dynamic field $v_t(\mathbf{r})$ using the basis function decomposition. The minimum distance between adjacent basis functions must also satisfy Shannon's sampling theorem to capture the dominant spectral information of the field. Thus, the minimum distance between basis functions must satisfy 
\begin{equation}\label{eq:BasisFunctionSeparation}
	\Delta_b < \frac{1}{2\rho_b\nu_{co}}. 
\end{equation}
where $\rho_b$ is an over-sampling parameter to determine the basis function separation and $\nu_{co}$ is the spatial cut-off frequency from the observation. For the Gaussian basis functions, $\phi(\mathbf{r})$, being considered, the basis function width parameter, $\sigma_{\phi}^2$, can also be inferred using spectral considerations~\cite{Sanner1992,Scerri2009}. The Fourier transform of an \textit{n}-dimensional Gaussian is another Gaussian given by
\begin{equation}\label{eq:GaussianFT}
\boldsymbol\Phi(\boldsymbol \nu)=\left(\frac{1}{\pi\sigma_{\nu}^2}\right)^{\frac{n}{2}}\mathrm{exp}\left(-\frac{1}{\sigma_{\nu}^2}\boldsymbol\nu^\top \boldsymbol\nu\right),
\end{equation}
where 
\begin{equation}\label{eq:GaussianFTWidth}
	\sigma^2_{\nu} = \frac{1}{\pi^2\sigma_{\phi}^2}. 
\end{equation}
The basis function width, $\sigma^2_{\nu}$, should be chosen to give at least 3 dB attenuation at $\boldsymbol\nu_{co}$ giving
\begin{equation}\label{eq:WidthFrequencyRelationship}
 \sigma^2_{\nu}=\frac{2\boldsymbol\nu_c^\top \boldsymbol\nu_c}{\ln 2}.
\end{equation}
This ensures that the basis functions can represent field with frequency content up to $\boldsymbol\nu_{co} $. Proofs for equations~\ref{eq:GaussianFT} and \ref{eq:WidthFrequencyRelationship} are given in~\ref{ap:FrequencyAnalysis}.
% Thus to ensure that the basis functions can represent fields with frequency content up to $\boldsymbol{\nu}_c$, $\sigma_{\nu}$ should be set with at least the 3~dB frequency for the basis functions equal to $\boldsymbol{\nu}_c$, thus obtained $\sigma^2_{\nu}$ and therefore $\sigma_{\phi}^2$.

>>>>

where $\mathbf{\boldsymbol{\phi}}(\mathbf{r})$ is a vector of Gaussian basis functions that are scaled by the state vector, $\mathbf{x}_t$. The choice of Gaussian basis functions can be justified by the existence of the so called bump solutions for this class of model, which have a Gaussian shape~\cite{Coombes2005}. The width and positioning of the basis functions can be determined by spectral analysis (explained in detail in Section~\ref{SpectralAnalysisSection}). The connectivity kernel can also be considered a decomposition into a set of basis functions where 

@@

There are two options available to alleviate problems with aliasing. The first, which is the most desirable, is to increase the spatial sampling frequency (using a higher resolution of sensors). The second is to use wider sensors, where the spatial averaging acts like a low-pass anti-aliasing filter. This option is a compromise between alleviating issues with aliasing and losing high frequency information contained in the field. In practice, it is difficult  to estimate the bandwidth of the cortex using traditional electrophysiological measurements due to spatial limitations. However, we envisage it may be possible the estimate it using other modalities with higher spatial resolution such as, fMRI, NIRS or other optical imaging techniques~\cite{Issa2000}. Errors associated with mild aliasing result in distortions in the high frequency components of the field dynamics. However, after applying the basis function decomposition, high frequency components contribute negligibly to the inverse problem and are therefore not a major problem for the state and parameter estimation~\cite{Sanner1992}. This is further illustrated our results. \omg{Although our results are severely effected by the loss of high frequency info! Ken: Do you think you can word this a little better than what I have done.}

Similar considerations need to be made regarding the representation of the dynamic field $v_t(\mathbf{r})$ using the basis function decomposition. The minimum distance between adjacent basis functions must also satisfy Shannon's sampling theorem to capture the dominant spectral information of the field. Thus, the minimum distance between basis functions must satisfy 
\begin{equation}\label{eq:BasisFunctionSeparation}
	\Delta_b < \frac{1}{2\rho_b\nu_{co}}. 
\end{equation}
where $\rho_b$ is an over-sampling parameter to determine the basis function separation and $\nu_{co}$ is the spatial cut-off frequency from the observations. For the Gaussian basis functions, $\phi(\mathbf{r})$, being considered, the basis function width parameter, $\sigma_{\phi}^2$, can also be inferred using spectral considerations~\cite{Sanner1992,Scerri2009}. The Fourier transform of an \textit{n}-dimensional Gaussian is another Gaussian given by
\begin{equation}\label{eq:GaussianFT}
\boldsymbol\Phi(\boldsymbol \nu)=\left(\frac{1}{\pi\sigma_{\nu}^2}\right)^{\frac{n}{2}}\mathrm{exp}\left(-\frac{1}{\sigma_{\nu}^2}\boldsymbol\nu^\top \boldsymbol\nu\right),
\end{equation}
where 
\begin{equation}\label{eq:GaussianFTWidth}
	\sigma^2_{\nu} = \frac{1}{\pi^2\sigma_{\phi}^2}. 
\end{equation}
The basis function width, $\sigma^2_{\nu}$, should be chosen to give at least 3 dB attenuation at $\boldsymbol\nu_{co}$ giving
\begin{equation}\label{eq:WidthFrequencyRelationship}
 \sigma^2_{\nu}=\frac{2\boldsymbol\nu_c^\top \boldsymbol\nu_c}{\ln 2}.
\end{equation}
This ensures that the basis functions can represent field with frequency content up to $\boldsymbol\nu_{co} $. Proofs for equations~\ref{eq:GaussianFT} and \ref{eq:WidthFrequencyRelationship} are given in~\ref{ap:FrequencyAnalysis}.
% Thus to ensure that the basis functions can represent fields with frequency content up to $\boldsymbol{\nu}_c$, $\sigma_{\nu}$ should be set with at least the 3~dB frequency for the basis functions equal to $\boldsymbol{\nu}_c$, thus obtained $\sigma^2_{\nu}$ and therefore $\sigma_{\phi}^2$.
